---
title: "Saxena_DSC520_Week11-12: 11.3"
author: "Amit Saxena"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

## Analysis on NFL Players- Week 8-9

### Introduction:

My son got into football when he started high school, and I started following 
football too. I’m an Indian and have no idea about this sport, but since my son 
started playing, I got interested in this sport and am trying to learn as much 
as possible about it. Since my son and his friends regularly talk about NFL, I 
thought this would be a great topic and a good opportunity for me to do some 
research on player profiles. 

### Research questions:

I want to research the data of NFL players and find out the traits which are 
common to all NFL players. I also want to compare how the players have evolved 
over the years by making comparisons between retired and active NFL players.

1.	What is the ideal physical profile to become an NFL player?
2.	Are there more chances of getting into NFL if you’re born in certain months 
    of the year?
3.	Does the geographical location matter? Meaning would an individual have a 
    better chance of getting into NFL, if he’s from a certain city or state?
4.  Does location matter in terms of the positions of the players? Meaning are 
    there more players from a certain state playing a particular position?
5.  Which college produces most NFL players?
6.	Which position has the tallest players?
7.  Which position weighs the most in average?
8.  How would the players look like in 10 years from now in terms of physical 
    profile?

### Approach:

*	I will clean up the file to remove the blank values and unnecessary data.
*	I’ll transform and modify the data.
*	I’ll separate the players by their positions.
*	Then I’ll summarize the results to find out the averages of the physical 
  parameters of the players.
*	I’ll try to use regression and predictions on the data.
*	I’ll plot the graphs and use scatter plots to see the trend.
*	Based on the inferences from the summary, plots and graphs I’ll try to draw 
  conclusions to the questions I have.

### How your approach addresses (fully or partially) the problem:

*	Using my approach, I’ll transform the data.
*	I’ll try to visualize the patterns of the graphs.
*	Based on the summarization, graphs and plots I’ll try to see the patterns to 
  answer the questions I have.
*	Based on the patterns, I’d try to arrive at some conclusions.

### Data:

I collected the data from NFL’s website as well as Kaggle, which some of the 
classmates suggested. 

The datasets I have chosen are:

1.	Active Players Profile
2.	Retired Players Profile
3.	Player Birth Dates
4.  Additional Player Data- hand & arm length (additional data collected in last
    2 weeks)

The links to the sources of data are:

1.	https://www.kaggle.com/datasets/trevyoungquist/2020-nfl-stats-active-and-retired-players
2.	https://www.nfl.com/players/
3.  https://nflbirthdays.com/

### Required Packages:

The packages I’m planning to use to begin with are, however as I start working 
on the project, I might find other packages to be used:

1.  dplyr: For data manipulation to solve the common data manipulations.
2.  readxl: For reading Excel files into R.
3.  utils: For utility functions that are useful in a variety of contexts.
4.  MASS: For performing a variety of statistical analyses, including linear and
    nonlinear regression, time series analysis etc.
5.  purrr: For functional programming toolkit for R to work more effectively 
    with functions and vectors.
6.  ggplot2: For creating graphics in R.
7.  readr: To read CSV files.
8.  Metrics: For metrics for evaluating machine learning models, including 
    regression models.
9.  pscl: For fitting and analyzing models for discrete data, such as count data
    and binary data.
10. caret: To train and evaluate a wide range of machine learning models.

```{r, include=FALSE}

library(dplyr)
library(readxl)
library(utils)
library(MASS)
library(purrr)
library(ggplot2)
library(readr)
library(Metrics)
library(pscl)
library(caret)
library(stringr)

```

```{r} 

# Importing the data from the downloaded files-

DS1 <- read_xlsx('/Users/asaxe/Documents/GitHub/DSC520-T303 Statistics for Data Science (2241-1)/DSC520- Final Project/NFLBirthdays_Copy.xlsx')
DS2 <- read_xlsx('/Users/asaxe/Documents/GitHub/DSC520-T303 Statistics for Data Science (2241-1)/DSC520- Final Project/Active_Player_Basic_Stats_Copy.xlsx')
DS3 <- read_xlsx('/Users/asaxe/Documents/GitHub/DSC520-T303 Statistics for Data Science (2241-1)/DSC520- Final Project/Retired_Player_Basic_Stats_Copy.xlsx')

# Displaying the data sample-
nrow(DS1)
head(DS1, 4)
nrow(DS2)
head(DS2, 4)
nrow(DS3)
head(DS3, 4)

```

### Plots and Table Needs:

I’m planning to use the following plots and tables, but as I progress, I may 
need to use some additional plots and tables:

1.	Scatterplot: I'll use it to see the data in a more graphical form to see the 
    trend over the years. For example how the height of quarterbacks.
2.	Histogram: I'll use this to see the counts for comparison. For ex. count of 
    hall of famers by the teams.
3.	Box Plot: I'll use box plot to see the distribution of a variable for ex. 
    month of birth of players.
4.	Line Plot: I'll use line plot to see the relationship between two continuous
    variables such as height and weight. 
5.	Bar Plot: I'll use bar plots the relationship between a categorical variable
    and a continuous variable, such as height against years.
6.  In terms of tables I'll use data frames, lists and matrices.


### Questions for future steps:

Some questions that I may include in future are:

1.	Is there really a pattern in NFL players?
2.	If there is a pattern then what could be the reason for such a pattern.
3.	Does the financial background of players is also a factor in them making it 
    to NFL?
4.	Are there certain states where NFL is more popular than other states?

Based on the approach and my analysis I may need to find some additional data.

#-----------------------------------------------------------------------------#
#-----------------------------------------------------------------------------#

##  Analysis on NFL Players- Week 10

###	How to import and clean my data

I imported the data I downloaded for this assignment using 'readxl' package, and 
diplayed the data last week.

This week I'll clean the data by removing the blank rows. Then I'll remove the 
values which are not valid, for ex. '--'. 
For the birthdays, I'll add columns for month and year of the birth to create 
new dataframes to help in visualization and sorting.
sort them 
based on the data type such as position etc.

```{r}

# Filtering out rows with missing data- 

DS_B <- DS1[!is.na(DS1$player),] # Filtered data set with player birthdays
DS_A <- DS2[!is.na(DS2$Player_Id),] # Filtered data set of active players 
DS_R <- DS3[!is.na(DS3$Player_Id),] # Filtered data set of retired players 

# Birthday dataset modifications-

# After filtering out the blank rows I noticed rows with '--' values, so filtering
## those as well in all data sets, just to be sure.
# Removing the rows which has '--' as the value
DS_B1 <- DS_B %>% filter(player != "--", position != "--", team != "--", 
                         birthday != "--") 
# Selecting the month from the birthday
DS_B1$birth_month <- substring(DS_B1$birthday, 6,7) 
# Selecting the year from the birthday
DS_B1$birth_year <- substring(DS_B1$birthday, 1,4) 
# Removing administrative positions such as coach, owner, manager etc.-
DS_B2 <- filter(DS_B1, !grepl('Coach|Team owner|Technical|Manager|Scout|
                              Supervisor|Official|NFL', position)) 

# Active player dataset modifications-

# Removing the rows which has '--' as the value
DS_A1 <- DS_A %>% filter(Player != "--", Position != "--", Height != "--", 
                         Weight != "--", College != "--", Number != "--", 
                         Current_Team != "--", Experience != "--", Age != "--") 
DS_A2 <- DS_A1[,-1] # Removing the first column as it's redundant
# Creating a column for height in inches by splitting the data & converting it to numeric-
DS_A2$Height_Inches <- (as.numeric(substring(DS_A2$Height, 6, 7)))*12 + 
  as.numeric(substring(DS_A2$Height, 9, 10))

# Retired player dataset modifications-

# Removing the rows which has '--' as the value
DS_R1 <- DS_R %>% filter(Full_Name != "--", Position != "--", Height != "--", 
                         Weight != "--", College != "--") 
DS_R2 <- DS_R1[,-1] # Removing the first column as it's redundant
# Creating a column for height in inches by splitting the data & converting it to numeric-
DS_R2$Height_Inches <- (as.numeric(substring(DS_R2$Height, 6, 7)))*12 + 
  as.numeric(substring(DS_R2$Height, 9, 10))
# Converting the format of weight-
DS_R2$Weight <- as.numeric(DS_R2$Weight)

```


###	What does the final data set look like?

```{r}
# Final dataset for further grouping and sorting-
head(DS_B2, 10) # Player birthday dataset
head(DS_A2, 10) # Active player dataset
head(DS_R2, 10) # Retired player dataset

summary(DS_B2)
summary(DS_A2)
summary(DS_R2)

```

###	Questions for future steps.

After I have the clean datasets to analyze, the questions for future steps would 
be-
1. How to club, summarize and visualize data to help get me the answers to my 
  questions.
2. How to best interpret the data, summary and visualization for better 
  understanding.
3. Is there some data missing due to which I might not get answers to some of 
  my questions.


###	What information is not self-evident?

From the raw data none of the information seems self evident to me. To get any 
information from the data we'll have to club, sort, summarize and visualize them.

###	What are different ways you could look at this data?

There are multiple ways to look at the data. We can look at it from the positions
and try to find the average height and weight of the players. 
We can also look at it by grouping the players by the colleges and try to find 
patterns there.
We can also look at it by the months the players are born in and try to find 
similarities there.

Visualizing birthday data by creating histograms-
```{r}

DS_B2 <- transform(DS_B2, birth_month = as.numeric(birth_month), 
                   birth_year = as.numeric(birth_year))
str(DS_B2)

par(mfrow = c(1,2))
hist(DS_B2$birth_month, xlim=c(1,12), ylim=c(0,150) , 
     main = "Players by Months (01-12)", xlab = "Months")
hist(DS_B2$birth_year,xlim=c(1890,2010),ylim=c(0,400) , 
     main = "Players by Decades", xlab = "Years(1880-2010)")
summary(DS_B2)

```
Visualizing and comparing Active and Retired player data by creating histograms, 
scatterplots and box plots-
```{r}
# Creating histograms-
par(mfrow = c(2,2),mar=c(6,3,2,1))
hist(DS_A2$Height_Inches, xlim=c(60,84), ylim=c(0,500) , 
     main = "Active Players by Height", xlab = "Height in Inches")
hist(DS_R2$Height_Inches, xlim=c(60,84), ylim=c(0,12500) , 
     main = "Retired Players by Height", xlab = "Height in Inches")
hist(DS_A2$Weight, xlim=c(150,400), ylim=c(0,500) , 
     main = "Active Players by Weight", xlab = "Height in Inches")
hist(DS_R2$Weight, xlim=c(150,400), ylim=c(0,10000) , 
     main = "Retired Players by Weight", xlab = "Height in Inches")

# Creating box plots-
par(mfrow = c(1,2),mar=c(6,3,2,1))
boxplot(Height_Inches ~ Weight, data = DS_A2, main="Active Players")
boxplot(Height_Inches ~ Weight, data = DS_R2, main="Retired Players")

#Creating scatterplots for weight and height of active and retired players-
par(mfrow = c(1,2))
attach(DS_A2)
plot(Weight, Height_Inches, main="Active Players",
   xlab="Weight ", ylab="Height", pch=19)

attach(DS_R2)
plot(Weight, Height_Inches, main="Retired Players",
   xlab="Weight ", ylab="Height", pch=19)

```

###	How do you plan to slice and dice the data?

- For next week, I'll focus more into position wise analysis of data, and try to 
get more insights into it.

###	How could you summarize your data to answer key questions?

- Right now, I don't think I can answer all questions but some of the trends are 
visible in the plots.

1.	What is the ideal physical profile to become an NFL player?

Ans.    I think it'd be better to answer this question next week when I have split 
        and analyzed the data by positions. Comparing an Offensive Lineman
        to a Wide Receiver doesn't seem fair.

2.	Are there more chances of getting into NFL if you’re born in certain months 
    of the year?

Ans. As per the data I have, January seems to be the month with most NFL players.

3.	Does the geographical location matter? Meaning would an individual have a 
    better chance of getting into NFL, if he’s from a certain city or state?

Ans. I need to have more data to accurately reflect the players location, instead
    of using the college.
    
4.  Does location matter in terms of the positions of the players? Meaning are 
    there more players from a certain state playing a particular position?

Ans. I'll have an answer to this if I can find out more data on loaction.

5.  Which college produces most NFL players?

Ans. I haven't arrived at this answer yet, as the no. of colleges is large. I'll
    need to think how to solve this problem.

6.	Which position has the tallest players?

Ans. I plan to analyze position specific data next week.

7.  Which position weighs the most in average?

Ans. I plan to analyze position specific data next week.

8.  How would the players look like in 10 years from now in terms of physical 
    profile?

Ans. I'm planning to create a dataframe to create a model and predict the height
    and weight of the future players.

#### Birthday Data results-
- It's evident from the month based histogram that a disproportionate no. of 
players are born in January.
- I'll need to explore more to find the reson behind that, as currently I don't
have the data to analyze that.
- The year based histogram shows that most of the players were born in 1990-2000,
but that may only be due to the data we have, and I don't read too much into it.

#### Active and Retired Player Data results- 
- The median and mean weight of active players is much higher than the retired
players, which means the size of players have increased over the years.
However, next week I'll explore the position specific weight distribution to 
check if this increase is in general or skewed towards certain positions.
Below is a summary of weight side by side-

**Weight- Active**
<br>  Min.   :160.0
<br>  1st Qu.:205.0
<br>  Median :235.0
<br>  Mean   :244.4
<br>  3rd Qu.:287.0
<br>  Max.   :380.0

**Weight- Retired**
<br>  Min.   : 119.0
<br>  1st Qu.:195.0
<br>  Median :220.0
<br>  Mean   :228.0
<br>  3rd Qu.:252.0
<br>  Max.   :388.0
  
- The median height of active and retired players is same, but the mean height 
is slightly more for active players. Again, I'll explore this data more by 
position to see the details.
Below is a summary of height in inches side by side-
  
**Height_Inches- Active**
<br>Min.   :66.00
<br>1st Qu.:73.00
<br>Median :74.00
<br>Mean   :74.21
<br>3rd Qu.:76.00
<br>Max.   :81.00

**Height_Inches- Retired**
<br>Min.   :61.00
<br>1st Qu.:73.00
<br>Median :74.00
<br>Mean   :73.58
<br>3rd Qu.:75.00
<br>Max.   :85.00

###	What types of plots and tables will help you to illustrate the findings to 
### your questions?

- I will try to use pie chart for percentage distribution but I'll need to work 
more on data frames to be able to achieve that.

###	Do you plan on incorporating any machine learning techniques to answer your 
### research questions? Explain.

- I want to create a model to train, test and predict the height and weight of 
players 10 and 20 years from now and see how the results come out.
- I might have to create some dummy data to be able to plug in the predictions,
but I'll see if I'm actually able to achieve that.
- I wanted to include a machine learning and prediction model, but I couldn't
think of a way to incorporate it in my project.
- My recommended approach for machine learning would be to analyze the increase 
in the average weight and average height of the players by position and year of 
birth, and build a prediction model to forecast what the average weight and 
average height of the players be in 10 years time.
- Another approach could be to analyze the count of players from different states, 
and see the trend. By studying this trend for the available data, I'd like to 
forecast the estimated percentage of players from each state in 10 years and 20 
years.

###	Questions for future steps.

- I need to research why most of the players are born in January.
- I need to research and visualize the data specific to hall of famers more.
- I need to find a way to find the data regarding the location of the players. 
- I initially thought I could get that from the college, but that is not what I
initially planned to find out. So I'll try to find that data or try to extract 
it from NFL.com.

#-----------------------------------------------------------------------------#
#-----------------------------------------------------------------------------#

##  Analysis on NFL Players- Week 11-12

###	Introduction

I wanted to research the data of NFL players and find out the traits which are 
common to all NFL players. I also wanted to compare how the players have evolved 
over the years by making comparisons between retired and active NFL players.

As part of the analysis of NFL players for the final part of this project, I 
extracted data from NFL's website to include the length of arms and hands  of 
the players. 

I'll now clean up, merge and try to slice and dice this new joint dataset to 
see the trends by the positions of the players.

```{r}

# Importing the new dataset with values of arms and hands of the active players
# that I extracted from NFL website-

DS4 <- read_xlsx('/Users/asaxe/Documents/GitHub/DSC520-T303 Statistics for Data Science (2241-1)/DSC520- Final Project/Active_Player_Additional_Data.xlsx')

# Displaying the new data-
nrow(DS4)
head(DS4, 4)

DS_HA <- DS4[!is.na(DS4$Arms),] # Filtering the data set

# Merging the additional data with the active players data-
DS_HA1 <- merge(DS_A2, DS_HA, by="Player") 
nrow(DS_HA1)
head(DS_HA1)

# Removing columns which are not required
DS_HA2 <- DS_HA1[ -c(3:4,6:13) ]
head(DS_HA2)

# Define the position names of the position-
pos_names = c('C'='Center','CB'='Corner Back','DB'='Defensive Back',
              'DE'='Defensive End','DT'='Defensive Tackle','FB'='Full Back',
              'FS'='Free Safety','ILB'='Inside Line Backer','K'='Kicker',
              'LB'='Line Backer','LS'='Long Snapper','MLB'='Middle Line Backer',
              'NT'='Nose Tackle','OG'='Offensive Guard','OL'='Offensive Line',
              'OLB'='Outside Line Backer','OT'='Offensive Tackle','P'='Punter',
              'QB'='Quarter Back','RB'='Running Back','SAF'='Safety',
              'SS'='Strong Safety','TE'='Tight End','WR'='Wide Receiver',
              'G'='Guard','T'='Tackle')

# Add a new column for position names and replace the position abbreviations 
# with position names-
DS_HA2$'PositionName' <- DS_HA2$Position
DS_HA2$PositionName <- str_replace_all(DS_HA2$PositionName, pos_names)

# Summarize data by position names-
DS_HA2 %>% group_by(PositionName) %>% summarize_all(mean)

DS_HA3 <- DS_HA2 %>% group_by(PositionName) %>% summarize_all(mean)
DS_HA4 <- DS_HA3[-c(2:4)]

DS_HA4$`Height in Inches` <- as.numeric(DS_HA4$`Height in Inches`)
DS_HA4$Weight.y <- as.numeric(DS_HA4$Weight.y)
DS_HA4$Arms <- as.numeric(DS_HA4$Arms)
DS_HA4$Hands <- as.numeric(DS_HA4$Hands)
DS_HA4

# Rename column names to better highlight the values-
colnames(DS_HA4) <- c('PositionName','Mean Height','Mean Weight','Mean Arm Length'
                      ,'Mean Hand Length')
print(DS_HA4)

# Create bar graphs for the different statistics against the position-
barplot(DS_HA4$`Mean Height`,names.arg=DS_HA4$PositionName,ylab="Mean Height", 
        col= 'skyblue', las=2, border = 0, cex.lab=1, cex.axis=1, font=1,col.axis="black")

barplot(DS_HA4$`Mean Weight`,names.arg=DS_HA4$PositionName,ylab="Mean Weight", 
        col= 'blue', las=2, border = 0, cex.lab=1, cex.axis=1, font=1,col.axis="black")

barplot(DS_HA4$`Mean Arm Length`,names.arg=DS_HA4$PositionName,ylab="Mean Arm Length", 
        col= 'lightblue', las=2, border = 0, cex.lab=1, cex.axis=1, font=1,col.axis="black")

barplot(DS_HA4$`Mean Hand Length`,names.arg=DS_HA4$PositionName,ylab="Mean Hand Length", 
        col= 'lightgreen', las=2, border = 0, cex.lab=1, cex.axis=1, font=1,col.axis="black")

```

```{r}
# Visualize college representation of the players-

DS_A2_C <- aggregate(Player ~ College, data = DS_A2, function(x) length(x))
DS_A2_C1 <- DS_A2_C[order(DS_A2_C$Player, decreasing = TRUE),] 
head(DS_A2_C1,10)

barplot(DS_A2_C1$Player,names.arg=DS_A2_C1$College,ylab="No. of Players", 
        col= 'lightgreen', las=2, border = 0, cex.lab=1, cex.axis=1, font=1,col.axis="black")

```


```{r}
# Analyzing the Hall of Famers by position-

# Expanding the position names (some names are old and not valid anymore)-
DS_R2$Position <- str_replace_all(DS_R2$Position, pos_names)
DS_R2_HoF <- aggregate(Hall_Of_Fame ~ Position, data = DS_R2, function(x) length(x))
DS_R2_HoF1 <- DS_R2_HoF[order(DS_R2_HoF$Hall_Of_Fame, decreasing = TRUE),] 
head(DS_R2_HoF1,10)

barplot(DS_R2_HoF1$Hall_Of_Fame,names.arg=DS_R2_HoF1$Position,ylab="No. of Hall of Famers", 
        col= 'lightgreen', las=2, border = 0, cex.lab=1, cex.axis=1, font=1,col.axis="black")

# Analyzing by College-
DS_R2_HoF_C <- aggregate(Hall_Of_Fame ~ College, data = DS_R2, function(x) length(x))
DS_R2_HoF_C1 <- DS_R2_HoF_C[order(DS_R2_HoF_C$Hall_Of_Fame, decreasing = TRUE),] 
head(DS_R2_HoF_C1,10)

#barplot(DS_R2_HoF_C1$Hall_Of_Fame,names.arg=DS_R2_HoF_C1$College,ylab="No. of Hall of Famers by College", 
#        col= 'lightblue', las=2, border = 0, cex.lab=1, cex.axis=1, font=1,col.axis="black")

```

###	The problem statement you addressed

- I started with a set of questions that I wanted to answer. 
- I answered some of them based on the analysis I did in week 10. However, there 
were several questions to which I had no answer and that needed additional 
analysis, which I didn't last and this week. 
- I have now found answers to almost all questions that I wanted to answer.

Below are some questions that I couldn't answer till last week, and attempeted
to answer this week based on additional data and analysis-

1. I wanted to research why most of the players are born in January.

2. I wanted to research and visualize the data specific to hall of famers more.

3. I wanted to find a way to find the data regarding the location of the players

4.  Which college produces most NFL players?

5.	Which position has the tallest players?

6.  Which position weighs the most in average?

###	How you addressed this problem statement

The past two weeks I carried out additional analysis based on the position of the
players, which gave me answers to my remaining questions. I also had to extract
additional data from NFL website to compare players.

I also did additional analysis on the hall of fame data for the retired players.

Here's the answers to the questions I couldn't answer in week 10-

1. I wanted to research why most of the players are born in January.

Ans 1. This to the best of my understanding is due to the cutoff being late in the 
    year for school admissions in US. Due to this the kids born in January have 
    a physical advantage over the kids in the same grade but born later in the 
    year.
    
    I believe this advantage carries on year after year and due to this comparative
    advantage more kids born in January tend to be selected for NFL.

2. I wanted to research and visualize the data specific to hall of famers more.

Ans 2. As per my data and analysis-

  The top 5 positions for Hall of Fame NFL players are:
  
      1. Defensive Back-	3465		
      2. Wide Receiver-	3025		
      3. Line Backer-	2953		
      4. Running Back-	2340		
      5. Guard-	2038	
      
  I was expecting QBs in the top 5 but surprisingly for me they are not even
  in top 10 in the list.
    
  The top 5 Colleges for Hall of Fame NFL players are:
  
      1. Notre Dame-	528		
      2. USC-	501		
      3. Ohio State-	432		
      4. Nebraska-	377		
      5. Penn State-	371	
      
  I was expecting colleges from Texas and Florida in this list, but surprisingly
    there is only 1 Florida college in 7th place.
      
3. I wanted to find a way to find the data regarding the location of the players. 

Ans 3. I initially thought I could get that from the college, but that is not 
      what I initially planned to find out.

Ans 3 (contd.). I couldn't find reliable data for sufficient no. of players to 
      arrive at any conclusion on this question. 
    
4.  Which college produces most NFL players?

Ans 4. Based on the data I collected following are the top 5 colleges
    represented by active NFL players-
    
      1.  Alabama	51		
      2.	Ohio State	42		
      3.	LSU	33		
      4.	Miami (Fla.)	32		
      5.	Florida	31	

5.	Which position has the tallest players?

Ans 5. As per the data the following are the positions that are tallest on average-

      1.  Offensive Tackle
      2.  Tight End
      3.  Offensive Guard
      4.  Defensive End
      5.  Nose Tackle

6.  Which position weighs the most in average?

Ans 6. As per the data the following are the positions that weigh most on average-

      1.  Nose Tackle
      2.  Offensive Tackle
      3.  Offensive Guard
      4.  Defensive Tackle
      5.  Center

###	Analysis

- I conducted analysis on 4 datasets that I collected. Those datasets were on active
 players, retired players, birthday of players and the length of hands ans arms 
 of the players.
- I cleaned up the data to remove the blank entries. Then I organized the data so 
 that it's easier to work with. I organized the columns, renamed them, removed 
 the ones I didnt need, added columns, renamed the positions, merged dataframes,
 and restructured the data.
- I created several bars and graphs to visualize the trends and tried to interpret
 them to answer some questions.
- I calculated the means of physical attributes of all active players and retired 
 players for comparison. Then I broke them into individual positions and then 
 calculated the means again for better understanding.
- I found the positions which have the heaviest players and positions which have 
 the tallest players in football.
- On retired players data, I segregated the data to find which positions have the 
 most no. of hall of famers, and which colleges have the most representation in
 hall of famers. I found some surprising facts there which I mentioned earlier.

###	Implications

- I found some very interesting information which contradicted my assumptions about 
the players, colleges and the positions, which I think is amazing, and what data 
science is supposed to do exactly. 
- I was assuming that a lot of quarterbacks get inducted to Hall of Fame because 
generally they're the stars of the teams, but that is not the case based on the
data I saw.
- I also assumed that colleges from Texas and Florida, being important football states, 
should have more Hall of Fame inductees, but again that is not true as per the
data I collected.
- A lot of times we make some stereotypes in our minds without any factual data, 
and having data science break those stereotypes is very important, 
based on the story the data tells.

- Another trend that I noticed is the height of active players doesn't seem to have
increased over the retired players, but the weight of players definitely seem to 
have increased.

###	Limitations

- The data I collected may not have been fully upto date and may have some errors 
in it, which I couldn't verify.
- I also couldn't find any data to analyze the locations of the players to have an
understanding of which states have the biggest contribution towards NFL players.

###	Concluding Remarks

- I thoroughly enjoyed working on this project and wish I had more time to further
analyze the data. I learned a lot of new things in R while working on this project
and I'm glad I picked the subject which I found interesting.
- This project also enlightened me on why Data Science is becoming such an important
area these days. The amount of data we generate can be utilized to understand a 
lot of things and based on that understanding so many improvements can be carried
out.
- I'd like to continue working on this topic and try to find more insights from the
data.
- I wanted to include a machine learning and prediction model, but due to lack of
time I couldn't incorporate it in my project.
- My recommended approach for machine learning would be:

  1. To analyze the increase in the average weight and average height of the 
  players by position and year of birth, and build a prediction model to forecast 
  what the average weight and average height of the players be in 10 years time.
  
  2. Another approach could be to analyze the count of players from different states, 
  and see the trend. By studying this trend for the available data, I'd like to 
  forecast the estimated percentage of players from each state in 10 years and 20 
  years.
